A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest}

### 2023-02-10 manuscript/manuscript.tex
fixed tables 3 decimals, expanded abstract

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest},
  pdfauthor={James M Brophy},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac
Arrest \\\large{A Bayesian Perspective} }
\author[1]{James M Brophy%
\corref{cor1}%
\fnref{fn1}}
 \ead{james.brophy@mcgill.ca}

@@

\begin{abstract}
A recent randomized clinical trial reported in patients with refractory
out-of-hospital cardiac arrest, extracorporeal CPR and conventional CPR
had similar effects on survival with a favorable neurologic outcome.
Herein, it is examined whether a Bayesian perspective allows any
additional insights into the interpretation of this trial.
\end{abstract}

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

This communication does not reiterate the many reasons to be wary of
null hypothesis significance testing (NHST), p values and confidence
intervals\citep{RN5420}. Rather it assumes the reader has perhaps heard
that Bayesian methods mirror our intuitive learning process and is
curious about its potential application to RCT interpretations.

@@

Posterior distributions are summarized with medians and 95\%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95\% of the probability density function
\citep{mcelreath2020}. We not only calculated the posterior probability
of any additional survival with eCPR (OR \textgreater1.00), but also of
clinically meaningful benefits (defined as OR \textgreater1.10).

ITT assesses subjects based on the group they were initially (and
randomly) allocated to, regardless of whether or not they dropped out,
were fully adhered to the treatment or switched to an alternative
treatment. ITT analyses can therefore be seen as a conservative estimate
which mirrors clinical effectiveness. In contrast, a per protocol (PP)
analysis involves a comparison of treatment groups in a trial that
includes only those patients who completed the treatment they were
originally allocated to. Similarly an ``as treated'' analysis considers
only which treatments subjects received, regardless of their
randomization status and protocol adherence. While both PP and
as-treated analyses alone may lead to bias, in conjunction with an ITT
analysis they may provide additional insights into efficacy and have
also been examined from a Bayesian perspective.

@@

ITT data from the INCEPTION trial\citep{CPR2023a} and two other
pertinent trials\citep{RN6751, RN6759} that also randomized CPR to eCPR
are shown in Table 1. Performing a Bayesian analysis on the INCEPTION
trial, using the default vague prior, produces an odds ratio (OR) 1.32
(95\% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result
to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms no impact
of the vague prior and shows this Bayesian analysis is completely
dominated by the observed INCEPTION data.

@@

\setlength{\LTpost}{0mm}
\begin{longtable}{lccccr}
\caption*{
{\large Table 2 eCPR odds ratios, 95\% credible intervals and probabilities with various priors}
} \\ 
\toprule
Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{2}{c}{Probabilities} \\ 
\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-6}
 & point estimate & lower limit & upper limit & p(OR) >1  &  p(ROPE) \\ 
\midrule
Vague & 1.32 & 0.543 & 3.21 & 0.727 & 0.134 \\ 
Combined & 1.35 & 0.705 & 2.58 & 0.817 & 0.153 \\ 
Enthusiastic & 1.40 & 0.738 & 2.67 & 0.849 & 0.138 \\ 
Skeptical & 1.32 & 0.700 & 2.50 & 0.804 & 0.168 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Vague: default vague prior\\
Combined: prior eCPR data from ARREST + PRAGUE\\
Enthusiastic: prior eCPR data from ARREST alone\\
Skeptical: prior eCPR data from PRAGUE alone\\
ROPE: range of practical equivalence = + / - 10\% OR (odds ratio)\\
\end{minipage}

\setlength{\LTpost}{0mm}
\begin{longtable}{lccccr}
\caption*{
{\large Table 3 eCPR (per protocol) odds ratios, 95\% credible intervals and probabilities with various priors}
} \\ 
\toprule
Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{2}{c}{Probabilities} \\ 
\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-6}
 & point estimate & lower limit & upper limit & p(OR) >1  &  p(ROPE) \\ 
\midrule
Vague & 0.451 & 0.151 & 1.35 & 0.0698 & 0.0473 \\ 
Combined & 0.859 & 0.430 & 1.71 & 0.33 & 0.2022 \\ 
Enthusiastic & 0.870 & 0.437 & 1.73 & 0.345 & 0.2013 \\ 
Skeptical & 0.858 & 0.419 & 1.75 & 0.327 & 0.1970 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Vague: default vague prior\\
Combined: prior eCPR data from ARREST + PRAGUE\\
Enthusiastic: prior eCPR data from ARREST alone\\
Skeptical: prior eCPR data from PRAGUE alone\\
ROPE: range of practical equivalence = + / - 10\% OR (odds ratio)\\
\end{minipage}

>>>>

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Bayesian Perspective},
  pdfauthor={James M Brophy MD PhD},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory
Out-of-Hospital Cardiac Arrest} }
\author[1]{James M Brophy MD PhD%
\corref{cor1}%
\fnref{fn1}}
 \ead{james.brophy@mcgill.ca}

@@

\begin{abstract}
A recent randomized clinical trial, INCEPTION, reported patients with
refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and
conventional CPR (cCPR) had similar effects on survival with a favorable
neurologic outcome. The current study examines if a Bayesian perspective
provides additional insights. Depending on the prior selected, the
Bayesian approach for the INCEPTION intention-to treat (ITT) analysis
shows an equivalence probability between 13.4 - 16.8\% (defined as 1 /
1.1 \textless{} odds ratio (OR) \textless{} 1.1). The probability of
clinical superiority with eCPR ranges from 65.7 - 77.0 \% (defined as OR
\textgreater{} 1.1). A similar analyses using INCEPTION per protocol
data shows an equivalence probability between 4.7 - 20.2\% with reduced
probabilities of clinical superiority of approximately 25\%. It is
concluded that a Bayesian perspective allows considerable additional
insights into the trial analysis and interpretation. The substantial
probabilities of increased survival, but with considerable residual
uncertainties, suggests that additional studies are required before
concluding that eCPR and cCPR have similar survival effects.
\end{abstract}

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

This communication does not reiterate the many reasons to be wary of
null hypothesis significance testing (NHST), p values and confidence
intervals\citep{RN5420}. Rather it assumes the reader has perhaps heard
that Bayesian methods mirror our intuitive learning and diagnostic
processes and is curious about its potential application to RCT analyses
and interpretations.

@@

Posterior distributions are summarized with medians and 95\%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95\% of the probability density function
\citep{mcelreath2020}. Bayesian analyses permit not only calculated the
posterior probability of any additional survival with eCPR (OR
\textgreater1.00), but also of clinically meaningful benefits. While
there is no universal standard for a clinically meaningful benefit, a
survival OR \textgreater1.10 may be an acceptable threshold for many.
Bayesian analyses also allows calculation of the probability between any
two points. For example, rather than simply comparing if the survival of
one treatment is better, one can calculate a range of practical
equivalence (ROPE). While different ranges may be proposed as
equivalent, +/- 10\% seems a reasonable starting point.

ITT assesses subjects based on the group they were initially (and
randomly) allocated to, regardless of whether or not they dropped out,
were fully adhered to the treatment or switched to an alternative
treatment. ITT analyses can therefore be seen as a conservative estimate
which mirrors clinical effectiveness. In contrast, a per protocol (PP)
analysis accounts for adherence by analyzing only those patients who
completed the treatment they were originally allocated to. While PP may
provide additional insights into efficacy, it is subject to bias. Since
a PP analysis when performed in conjunction with an ITT analysis may
provide additional insights, this has also been subjected to a Bayesian
analysis.

@@

ITT data from the INCEPTION trial\citep{CPR2023a} and two other
pertinent trials\citep{RN6751, RN6759} that also randomized CPR to eCPR
are shown in Table 1. Performing a Bayesian analysis on the INCEPTION
trial, using the default vague priors, produces an odds ratio (OR) 1.32
(95\% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result
to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms minimal
impact of the vague prior and shows this Bayesian analysis is completely
dominated by the observed INCEPTION data.

@@

\setlength{\LTpost}{0mm}
\begin{longtable}{lcccccc}
\caption*{
{\large Table 2 eCPR odds ratios, 95\% credible intervals and probabilities with various priors}
} \\ 
\toprule
Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{3}{c}{Probabilities} \\ 
\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-7}
 & point estimate & lower limit & upper limit & p(OR) >1  & p(OR) >1.1  &  p(ROPE) \\ 
\midrule
Vague & $1.321$ & $0.543$ & $3.215$ & 0.727 & $0.657$ & 0.134 \\ 
Combined & $1.349$ & $0.705$ & $2.580$ & 0.817 & $0.732$ & 0.153 \\ 
Enthusiastic & $1.403$ & $0.738$ & $2.668$ & 0.849 & $0.770$ & 0.138 \\ 
Skeptical & $1.322$ & $0.700$ & $2.496$ & 0.804 & $0.710$ & 0.168 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Vague: default vague prior\\
Combined: prior eCPR data from ARREST + PRAGUE\\
Enthusiastic: prior eCPR data from ARREST alone\\
Skeptical: prior eCPR data from PRAGUE alone\\
ROPE: range of practical equivalence = + / - 10\% OR (odds ratio)\\
\end{minipage}

\setlength{\LTpost}{0mm}
\begin{longtable}{lcccccc}
\caption*{
{\large Table 3 eCPR (per protocol) odds ratios, 95\% credible intervals and probabilities with various priors}
} \\ 
\toprule
Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{3}{c}{Probabilities} \\ 
\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-7}
 & point estimate & lower limit & upper limit & p(OR) >1  & p(OR) >1.1  &  p(ROPE) \\ 
\midrule
Vague & $0.451$ & $0.151$ & $1.348$ & $0.070$ & $0.049$ & $0.047$ \\ 
Combined & $0.859$ & $0.430$ & $1.713$ & $0.330$ & $0.238$ & $0.202$ \\ 
Enthusiastic & $0.870$ & $0.437$ & $1.734$ & $0.345$ & $0.251$ & $0.201$ \\ 
Skeptical & $0.858$ & $0.419$ & $1.755$ & $0.327$ & $0.237$ & $0.197$ \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Vague: default vague prior\\
Combined: prior eCPR data from ARREST + PRAGUE\\
Enthusiastic: prior eCPR data from ARREST alone\\
Skeptical: prior eCPR data from PRAGUE alone\\
ROPE: range of practical equivalence = + / - 10\% OR (odds ratio)\\
\end{minipage}
